The Effect of a Second AIDS Defining Illness on Survival in the MACS

### 2015-05-03 paper.Rmd
Work on naive PH Model

```{r chunks, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

We read in the data.
```{r data}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

# read in horiz data with lots of covariates
covariates <- read_csv("data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           age=as.Date("1960-01-01") + DOB,
           exit=ifelse(!died, as.Date("1994-12-31"), death.date),
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           dead.at.censor=(death.date < as.Date("1994-12-31"))) %>%
    select(STUDYID, frst.aids.diag, censor.date, sec.aids.diag,
           dead.at.censor)

@@

    # keep MACS only
    filter(COHORT==1,
           # keep people with first ADI only
           frst.aids.diag != 0,
           # restrict to haart era visits right now
           haart.era.visit == TRUE) %>%
    mutate(censor.date=as.Date("1970-01-01")+censor.date) %>%
    select(id=STUDYID, center=CENTER, visit:dead.at.censor)
```

@@

theme_set(theme_classic())

visits %>%
    group_by(id) %>%
    summarise(sec.aids=min(sec.aids.diag)) %>%
    mutate(sec.aids=plyr::revalue(as.character(sec.aids),
                                  c("0"="No Aids",
                                    "1"="PCP",
                                    "2"="KS",
                                    "3"="NHL",
                                    "4"="Neuro",
                                    "5"="Bacteria OI",
                                    "6"="Fungal OI",
                                    "7"="Protozoa",
                                    "8"="CMV/HSV",
                                    "9"="Wasting",
                                    "10"="Recurrent PNA",
                                    "11"="Cervical Ca"))) %>%
ggplot(., aes(x=sec.aids)) + geom_histogram() +
    xlab("Type of Secondary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Figure 1") +
    theme(axis.text.x = element_text(angle=-45))
```

>>>>

```{r chunks, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      results='asis')
```

```{r libs}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(survival)
theme_set(theme_classic())
```

We read in the data.
```{r data}
# read in horiz data with lots of covariates
covariates <- read_csv("../data/macswihs_horiz2013_copy-1.csv") %>%
           # date of death
    mutate(death.date=as.Date("1960-01-01") + FRSTDTHD,
           # date of aids diagnosis
           aids.date=as.Date("1960-01-01") + FRSTAIDD,
           # date of secondary aids diagnosis
           sec.aids.date=as.Date("1960-01-01") + SECAIDD,
           # what was the first aids diagnosis
           frst.aids.diag=AIDSDIAG,
           # what was the second aids diagnosis
           sec.aids.diag=SECAIDSDIAG,
           # did the person have a sec aids diag
           had.sec=year(sec.aids.date) < 2100,
           # did the person die
           died=year(death.date) < 2100,
           age=as.Date("1960-01-01") + DOB,
           exit=ifelse(!died, as.Date("1994-12-31"), death.date),
           censor.date=as.Date("1970-01-01") + as.integer(exit),
           last.alive=as.Date("1960-01-01") + LASTALID,
           censor.date=ifelse(last.alive < censor.date, last.alive, 
                              censor.date),
           censor.date=as.Date("1970-01-01")+censor.date,
           dead.at.censor=(death.date < as.Date("1994-12-31"))) %>%
    select(STUDYID, 
           aids.date,
           sec.aids.date,
           frst.aids.diag, 
           censor.date, 
           sec.aids.diag,
           dead.at.censor, 
           death.date) %>%
    filter(year(aids.date) < 2100)

@@

    # keep MACS only
    filter(COHORT==1,
           # keep people with first ADI only
           frst.aids.diag != 0,
           # restrict to haart era visits right now
           haart.era.visit == TRUE) %>%
    select(id=STUDYID, center=CENTER, visit:dead.at.censor)
```

@@

data.plot <- visits %>%
    group_by(id) %>%
    summarise(sec.aids=min(sec.aids.diag)) %>%
    mutate(sec.aids=plyr::revalue(as.character(sec.aids),
                                  c("0"="No Aids",
                                    "1"="PCP",
                                    "2"="KS",
                                    "3"="NHL",
                                    "4"="Neuro",
                                    "5"="Bacteria OI",
                                    "6"="Fungal OI",
                                    "7"="Protozoa",
                                    "8"="CMV/HSV",
                                    "9"="Wasting",
                                    "10"="Recurrent PNA",
                                    "11"="Cervical Ca"),
                                  warn_missing=FALSE))
ggplot(data.plot, aes(x=sec.aids)) + 
    geom_histogram() +
    xlab("Type of Secondary Aids Diagnosis") +
    ylab("Count") +
    ggtitle("Figure 1") +
    scale_x_discrete(limits=names(sort(-table(data.plot$sec.aids)))) +
    theme(axis.text.x = element_text(angle=-90))
```

We fit a Cox proportional hazards model to the pre haart era data. We did not allow secondary aids diagnosis to be time varying. 
```{r naive-survival}
# http://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf
# https://stat.ethz.ch/R-manual/R-devel/library/survival/html/coxph.html
data <- covariates %>%
    filter(year(aids.date) <= 1994) %>%
    rowwise() %>%
    transmute(start=aids.date,
              stop=min(censor.date, as.Date("1994-12-31")),
              status=(death.date <= as.Date("1994-12-31")),
              sec.aids=(sec.aids.diag > 0)) %>%
    mutate_each(funs(as.numeric), start, stop) %>%
    filter(stop > start)

naive.model <- coxph(Surv(start, stop, status) ~ sec.aids,
                     data=data)
knitr::kable(coef(summary(naive.model)),
             caption="Naive Model")
```

```{r strat-surv}
# strat.model <- coxph(Surv(first.aids.date, stop, dead.at.censor) ~ 
#                      had.sec.aids + strata(haart.era),
#                      data=data)
```

```{r propensity-score}
# fit logit model (and nonparam?) to calculate prop score
```

```{r matching}
```
